---
title: "amelpool_fst"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{amelpool_fst}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(amel.poolseq)
library(ggplot2)
library(dplyr)
library(reshape2)
library(utils)
```

# Vignette for FST figures

## Define Variables
### From Raw
```{r eval=FALSE}
df <- amel_colnames(pop2names, popool2_fst)

samp_pop <- function(popnames)
  pop2_samp <- data.frame(t(combn(popnames[,1],2)))
  samp_name <- paste(pop2_samp$X1, pop2_samp$X2, sep='.')
  pop2_pop <- data.frame(t(combn(popnames[,2],2)))
  pop_name <- paste(pop2_pop$X1, pop2_pop$X2, sep='')
  amel_fst_df <- data.frame(cbind(samp_name, pop_name))
}

df_stats <- samp_pop(popnames)
````

### From Cluster
```
df <- read.table("all2.fst", header=T)
rm_chars <- function(x){
  as.numeric(gsub('.*_', '', x))
}
df <- df %>% mutate(across(6:ncol(df), rm_chars))

df_stats <- read.table("comp", header=F)
df_stats <- df_stats[,c(2,1)]
```

## Avg. FST per population
mean.n0 excludes zeroes.

```{r}
mean <- colMeans(df[6:ncol(df)])
mean.n0 <- apply(df[6:ncol(df)],2,function(x) mean(x[x>0]))
df_means <- cbind(mean, mean.n0, df_stats)
colnames(df_means) <- c("FST", "FST_n0", "stock", "comp")
rownames(df_means) <- NULL
```

## Aggregate avg. FST by management

```{r}
df_means <- subset(df_means, select=-c(1))
df_means2 <- df_means[order(df_means$FST_n0),]
df_agman <- df_means2 %>% group_by(comp) %>%
  summarize(FST=mean(FST_n0), sd=sd(FST_n0))
```

### Figure
```{r echo=FALSE}
ampop_avfst <- ggplot(df_means2, aes(x=comp, y=FST_n0, color=comp)) + geom_boxplot() + theme(legend.position="none",
        axis.title.x=element_text(color="black", size=15),
        axis.title.y=element_text(color="black", size=15)) +
  ylab("Average FST") + xlab("Management") + 
  ggtitle("Average FST by Comparison") + theme(plot.title=element_text(hjust=0.5))
ampop_avfst
```

### AOV
```{r}
df_aov <- aov(df_means2$FST_n0~df_means2$comp)
summary(df_aov)
df_tukey <- TukeyHSD(df_aov)
plot(df_tukey, las=1)
```

## Fst heatmap by stock
```{r}

df_hm <- df_means %>% mutate(
  Field1 = gsub('[.].*','', df_means$stock), 
  Field2 = gsub('.*[.]','', df_means$stock)
  ) %>% select(!(stock))

df_hm2 <- df_hm %>% relocate(Field2, .before=Field1) %>%
  rename(
    Field1=3,
    Field2=4) %>%
  bind_rows(df_hm)

##note: code to reorder columns by name (alphabetical)

ggplot(df_hm2, aes(Field1, Field2, fill=FST_n0)) + geom_tile() +
  scale_fill_gradient(low="#D4AF37", high="black") +
  geom_text(aes(label=sprintf("%.3f", FST_n0)), color="white", size=4) +
  labs(x=NULL, y=NULL) +
  coord_fixed()
```

## Manhattan plot
The first line of code removes unclassified sections of the honey bee genome for easier plotting.

```{r}
ampop_mhdf <- df[-grep("NW_", df$RefContig),]

ampop_mhdf <- ampop_mhdf %>%
  group_by(RefContig) %>%
  summarize(chr_len=max(WindowPos)) %>%
  mutate(tot=cumsum(chr_len)-chr_len) %>%
  select(-chr_len) %>%
  left_join(ampop_mhdf, ., by=c("RefContig"="RefContig")) %>%
  arrange(RefContig, WindowPos) %>%
  mutate(BPcum=WindowPos+tot)

axisdf = ampop_mhdf %>% group_by(RefContig) %>% summarize(center=( max(BPcum) + min(BPcum) ) / 2 )
```
Now that the dataframe has been assembled, the Manhattan plot can be constructed.

```{r}
ampop_manplot <- ggplot(ampop_mhdf, aes(x=BPcum, y=ampop_mhdf[,7])) +
  geom_point( aes(color=as.factor(RefContig)), alpha=0.8, size=1.3) +
  scale_color_manual(values = rep(c("grey", "skyblue"), 70 )) +
  scale_x_continuous( label = axisdf$RefContig, breaks= axisdf$center ) +
  scale_y_continuous() +
  theme_bw() +
  theme( 
    legend.position="none",
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

ampop_manplot + facet_wrap(~RefContig, scales="free")
```
