---
title: "amelpool_fst"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{amelpool_fst}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(amel.poolseq)
library(ggplot2)
library(dplyr)
```

# Vignette for FST figures

## Step 1: Defining Variables
```{r eval=FALSE}
df <- amel_colnames(pop2names, popool2_fst)
```

## Step 2: Avg. FST per Population
mean.n0 excludes zeroes.

```{r}
mean <- colMeans(df[[1]])
mean.n0 <- apply(df,2,function(x) mean(x[x>0]))
df_means <- cbind(mean, mean.n0, df[[2]])
colnames(df_means) <- c("FST", "FST_n0", "comp")
df_means
```

## Step 3: Aggregate mean FST by comparison class and plot-----

```{r}
df_means <- subset(df_means, select=-c(1))
df_means$stock <- rownames(df_means)
rownames(df_means) <- NULL
df_means2 <- df_means[order(df_means$FST_n0),]
df_fstag <- aggregate(df_means2$FST_n0, by=list(df_means2$comp), function(x) mean(x, na.rm=TRUE))
colnames(df_fstag) <- c("Comparison", "FST")
```

### Figure
```{r echo=FALSE}
ampop_avfst <- ggplot(data=df_fstag, aes(x=Comparison, y=FST, fill=Comparison)) + 
  geom_bar(stat="identity") + 
  theme_classic() +
  theme(legend.position="none",
        axis.title.x=element_text(color="black", size=15),
        axis.title.y=element_text(color="black", size=15)) +
  ylab("Average FST") + xlab("Stock") + 
  ggtitle("Average FST by Comparison") + theme(plot.title=element_text(hjust=0.5))
ampop_avfst <- ampop_avfst + scale_color_manual(values=c("#4D1434", "#903163", "#969FA7"))
ampop_avfst
```

### AOV
```{r}
df_means <- df_means[-grep("Ref", df$Field1),]
df_aov <- aov(df_means$FST_n0~df_means$comp)
summary(df_aov)
df_tukey <- TukeyHSD(df_aov)
plot(df_tukey, las=1)
```

## Fst Heatmap
```{r}
mean.n0 <- data.frame(apply(df,2,function(x) mean(x[x>0])))
mean.n0$names <- rownames(mean.n0)
rownames(mean.n0) <- NULL
colnames(mean.n0) <- c("FST", "Comp")
mean.n0$Field1 <- gsub('[.].*','', mean.n0$Comp)
mean.n0$Field2 <- gsub('.*[.]','', mean.n0$Comp)
mean.n0_2 <- subset(mean.n0, select=-c(2))

ggplot(mean.n0_2, aes(Field1, Field2, fill=FST)) + geom_tile() + 
  geom_text(aes(label=sprintf("%.3f", FST)), color="white", size=4) +
  labs(x=NULL, y=NULL) +
  coord_fixed() + theme_bw()
```

## Manhattan plot

The first line of code removes unclassified sections of the honey bee genome for easier plotting.
```{r}
ampop_mhdf <- df[-grep("NW_", df$RefContig),]
ampop_mhdf <- ampop_mhdf %>%
  group_by(RefContig) %>%
  summarize(chr_len=max(WindowPos)) %>%
  mutate(tot=cumsum(chr_len)-chr_len) %>%
  select(-chr_len) %>%
  left_join(ampop_mhdf, ., by=c("RefContig"="RefContig")) %>%
  arrange(RefContig, WindowPos) %>%
  mutate(BPcum=WindowPos+tot)

axisdf = ampop_mhdf %>% group_by(RefContig) %>% summarize(center=( max(BPcum) + min(BPcum) ) / 2 )
```
Now that the dataframe has been assembled, the Manhattan plot can be constructed.

```{r}
ampop_manplot <- ggplot(ampop_mhdf, aes(x=BPcum, y=df4[,7])) +
  geom_point( aes(color=as.factor(RefContig)), alpha=0.8, size=1.3) +
  scale_color_manual(values = rep(c("grey", "skyblue"), 70 )) +
  scale_x_continuous( label = axisdf$RefContig, breaks= axisdf$center ) +
  scale_y_continuous() +
  theme_bw() +
  theme( 
    legend.position="none",
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

ampop_manplot + facet_wrap(~RefContig, scales="free")
```